Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-16 paper.tex
update

\begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=2.5em, column sep=3.7em]
    { 
        &  &  &  & 
      \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
      \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
      \begin{smallmatrix} \text{disentangled}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
      \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
        &
      %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\ 
    };
  {
    [start chain] \chainin (m-2-4);
    \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
    \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
    \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
  }
  { 
    [start chain] \chainin (m-2-1);
    \chainin (m-2-2) [join={node[above,labeled] {\text{fusion}}}];
    \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
    \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
    \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
  }
  { 
    [start chain] 
    \chainin (m-3-1);
    { [start branch=A] \chainin (m-2-1)
      [join={node[left, labeled] {\alpha}}];
    }
  }
    %\chainin (m-2-2);
    %\chainin (m-2-3);
    %\chainin (m-2-4);
  {
    [start chain] 
    \chainin (m-3-5); 
    { [start branch=B] \chainin (m-2-5)
      [join={node[right, labeled] {\alpha}}];
    }
  }
\end{tikzpicture}

@@

It is worth noting that the continuation class @Kont@ is defined as a list of frames.
The frame class @Frame@ stores the information of a single call-site, 
i.e., the information that can be used to resume the interrupted computation.
A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)

@@

Similar to CESK machines, the machine state of AAM has a control expression,
an environment, a store and continuation, and in addition with a time-stamp.
But there are several notable differences between AAM's store and CESK machine's store.
In AAM, the store maps addresses to sets of values, so it stores all possible values
for a particular address. Thus dereferencing addresses becomes non-deterministic.
Also, the store performs joining \todo{?} rather than overwriting when updating elements of the store.
Third, the continuations are likewise allocated on store instead of formed a linked list,
and the continuation in a state then turns into a continuation address.

@@

The structure of function @step@ is similar to the concrete CESK machine, 
except non-determinism which makes @step@ return a list of reachable successor states.
In the beginning of @step@, we call function @tick@ to update the time.
Then we have two cases to consider: 1) The current control expression is @Let@,
and then the result of @App(f, ae)@ will be bound to variable @x@.
Firstly we retrieve the set of closures that @f@ may represents.
For each closure in the set, we do almost the same as what we did in the concrete 
CESK machines. But note that since the continuation is allocated on store (@kstore@),
a new continuation address @new_kaddr@ is constructed, and a new frame
@Frame(x, e, env)@ paired with the current continuation address @kaddr@
is joined into @new_kaddr@. Finally, a list of successors states is generated.

@@

\begin{lstlisting}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) 
      if isAtomic(ae) =>
      val closures = atomicEval(f, env, bstore).toList
      for (Clos(Lam(v, body), env_c) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
        val new_kaddr = allocKont(body)
        val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
        State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
      }
    case State(ae, env, bstore, kstore, kaddr, time) 
      if isAtomic(ae) =>
      val conts = kstore(kaddr).toList
      for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
        val baddr = allocBind(x, new_time)
        val new_env = env_f + (x -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(e, new_env, new_store, kstore, f_kaddr, new_time)
      }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
        val frame = Frame(x, e, env)
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(body, new_env, new_store, frame::konts, new_time)
      }
    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
      val Frame(x, e, env_f)::konts = konts
      val baddr = allocBind(x, new_time)
      val new_env = env_f + (x -> baddr)
      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
      List(State(e, new_env, new_store, konts, new_time))
  }
}
\end{lstlisting}

In the first case of pattern matching, we may have multiple choices of closure for callee @f@.
For each closure in the set, a new frame is constructed and pushed onto the stack.
The code for handling the second case is the same as the concrete CESK machines.
The function @drive@ is also updated (but not shown here) to prevent passing a state with
empty frame list into function @step@.

@@

@@

In last section, we showed that by keeping an unabstracted stack in the state 
space we can recover the precise call/return match. In this section, we start
our first step of transformation: \textit{fusing} an AAM with unbounded-stack.

In fact, as later we will eliminate the stack and transform it to continuations
that are represented by higher-order functions, it does not matter what kind of 
AAM we start from. 
Because transforming to higher-order functions forces us to sequentialize 
the order of abstract evaluation.
We choose to start from AAM with unbounded-stack just because it has a 
straightforward correspondence of pushdown semantics to the refunctionalized AAM.

\begin{lstlisting}
def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
  case Nil => seen
  case hd::tl =>
    if (seen.contains(hd)) return drive_step(tl, seen)

    val new_time = tick(hd)
    val new_states = hd match {
      case State(Let(x, App(f, ae), e), env, bstore, konts, time) 
        if isAtomic(ae) =>
        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
          val frame = Frame(x, e, env)
          val baddr = allocBind(v, new_time)
          val new_env = env_c + (v -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(body, new_env, new_store, frame::konts, new_time)
        }
      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
        konts match {
          case Nil => List()
          case Frame(x, e, f_env)::konts =>
            val baddr = allocBind(x, new_time)
            val new_env = f_env + (x -> baddr)
            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
            List(State(e, new_env, new_store, konts, new_time))
        }
    }
    drive_step(new_states ++ tl, seen + hd)
}

def analyze(e: Expr): Set[State] = drive_step(List(inject(e)), Set())
\end{lstlisting}

@@

for plain direct-style, the continue function handles more cases.
\todo{a defunctionalized object for non-determinism}

\begin{lstlisting}
def aval(state: State, todo: List[State], seen: Set[State]): Set[State] = { 
  if (seen.contains(state)) return collect(todo, seen)

  val newTime = tick(state)
  state match {
    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
      val newStates = for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
        val frame = Frame(x, e, env)
        val baddr = allocBind(v, newTime)
        val newEnv = c_env + (v -> baddr)
        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(body, newEnv, newStore, frame::konts, newTime)
      }   
      collect(newStates ++ todo, seen + state)
    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
      continue(konts, atomicEval(ae, env, bstore), bstore, newTime, todo, seen + state)
  }   
}

def collect(todo: List[State], seen: Set[State]): Set[State] = { 
  todo match {
    case Nil => seen
    case hd::tl => aval(hd, tl, seen)
  } 
}

def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time,
              todo: List[State], seen: Set[State]): Set[State] = konts match {
  case Nil => collect(todo, seen)
  case Frame(x, e, f_env)::konts =>
    val baddr = allocBind(x, time)
    val newEnv = f_env + (x -> baddr)
    val newStore = bstore.update(baddr, ds) 
    val newState = State(e, newEnv, newStore, konts, time)
    collect(newState::todo, seen)
}

def analyze(e: Expr): Set[State] = aval(inject(e), List(), Set()) 
\end{lstlisting}

@@

@@

TODO: summarize PDCFA, AAC

>>>>

\begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=2.5em, column sep=3.7em]
    { 
        &  &  &  & 
      \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
      \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
      \begin{smallmatrix} \text{disentangled}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
      \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
        &
      %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
        &
      %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
      \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\ 
    };
  {
    [start chain] \chainin (m-2-4);
    \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
    \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
    \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
  }
  { 
    [start chain] \chainin (m-2-1);
    \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearizing} \\ \text{+} \\ \text{fusing} \end{smallmatrix}}     }];
    \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
    \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
    \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
  }
  { 
    [start chain] 
    \chainin (m-3-1);
    { [start branch=A] \chainin (m-2-1)
      [join={node[left, labeled] {\alpha}}];
    }
  }
    %\chainin (m-2-2);
    %\chainin (m-2-3);
    %\chainin (m-2-4);
  {
    [start chain] 
    \chainin (m-3-5); 
    { [start branch=B] \chainin (m-2-5)
      [join={node[right, labeled] {\alpha}}];
    }
  }
\end{tikzpicture}

@@

It is worth noting that the continuation class @Kont@ is defined as a list of
frames.
The frame can be considered as evaluation context in reduction semantics.
The frame class @Frame@ stores the information of a single call-site, 
i.e., the information that can be used to resume the interrupted computation.
A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)

@@

Similar to CESK machines, the machine state of AAM has a control expression,
an environment, a store and continuation, and in addition with a time-stamp.
But there are several notable differences between AAM's store and CESK machine's store.
In AAM, the store maps addresses to sets of values, so it stores all possible values
for a particular address. Thus dereferencing addresses becomes non-deterministic.
Also, the store performs joining rather than overwriting when updating elements
of the store.
Third, the continuations are likewise allocated on store instead of formed a linked list,
and the continuation in a state then turns into a continuation address.

@@

The structure of function @step@ is similar to the concrete CESK machine, 
except non-determinism which makes @step@ return a list of reachable successor states.
In the beginning of @step@, we call function @tick@ to update the time.
Then we have two cases to consider:

1) The current control expression is @Let@,
and then the result of @App(f, ae)@ will be bound to variable @x@.
Firstly we retrieve the set of closures that @f@ may represent.
For each closure in the set, we do almost the same as what we did in the concrete 
CESK machines. But note that since the continuation is allocated on store (@kstore@),
a new continuation address @new_kaddr@ is constructed, and a new frame
@Frame(x, e, env)@ paired with the current continuation address @kaddr@
is joined into @new_kaddr@. Finally, a list of successors states is generated.

@@

\begin{lstlisting}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>
      val closures = atomicEval(f, env, bstore).toList
      for (Clos(Lam(v, body), env_c) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
        val new_kaddr = allocKont(body)
        val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
        State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
      }
    case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
      val conts = kstore(kaddr).toList
      for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
        val baddr = allocBind(x, new_time)
        val new_env = env_f + (x -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(e, new_env, new_store, kstore, f_kaddr, new_time)
      }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
        val frame = Frame(x, e, env)
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(body, new_env, new_store, frame::konts, new_time)
      }
    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
      konts match {
        case Nil => List()
        case Frame(x, e, env_f)::konts =>
          val baddr = allocBind(x, new_time)
          val new_env = env_f + (x -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          List(State(e, new_env, new_store, konts, new_time))
      }
  }
}
\end{lstlisting}

In the first case of pattern matching, we may have multiple choices of closure for callee @f@.
For each closure in the set, a new frame is constructed and pushed onto the stack.
The code for handling the second case is the same as the concrete CESK machines.

@@

\section{Linearization} \label{linear}

In last section, we showed that by keeping an unabstracted stack in the state 
space we can recover the precise call/return match.
From this section, we start describing the transformations step by step.
Our base machine is the AAM with unbounded stack.
In fact, as later we will transform the stack to higher-order functions
represented continuations, it does not matter what kind of AAM we start from. 
Because transforming to higher-order functions forces us to sequentialize 
the order of abstract evaluation, and neither need to construct a frame of list
nor to allocate continuations in store.
We choose to start from AAM with unbounded stack just because it has a 
equivalence of stack model to the final abstract definitional interpreters.

\todo{Unbounded Stack AAM is PD, justify this}

In Danvy's paper \textit{Defunctionalized Interpreters for Programming
  Languages}, he metioned for deterministic languages, a reduction semantics
is a structural operational semantics in continuation style, where the reduction
context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
Here we are facing a problem:
the underlying semantics of AAM is fundamentally non-deterministic.
But the evaluation context (the @Frame@ in our program) is not
non-deterministic, and the work list @todo@ actually implicitly handles the non-determinism,
if we simply refunctionalize the frames to functions, it would not helps toward
abstract definitional interpreters.

In order to address this problem, our first step of tramsformations
is to linearize all the non-deterministic choices.
This step turns the @step@ function to be deterministic.
Likewise the @Frame@ saves the information of the caller in concrete executions,
we introduce another layer of controls, and define a case class @NDCont@
that saves the information at a fork point when we have multiple target closures.
We also add a new field @ndk@ to the definition of state which is called
@NDState@ now.

\begin{lstlisting}
case class NDCont(cls: List[Clos], argvs: Set[Storable], store: BStore, time: Time, frames: List[Frame])
case class NDState(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time, ndk: List[NDCont]) {
  def toState: State = State(e, env, bstore, konts, time)
}
\end{lstlisting}

A @NDCont@ contains a list of closures that are possible functions to be invoked,
a set of values that will be bind to the function's formal argument,
and the store, time, and frames at the fork point.
In the definition of @NDState@, an auxiliary function @toState@ is added that
converts itself to @State@.

Now the type of function @step@ becomes @NDState => NDState@.
The following code shows the first case of matching an instance of @NDState@ in
the @step@.

\begin{lstlisting}
case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
  val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
  val Clos(Lam(v, body), c_env) = closures.head
  val frame = Frame(x, e, env)
  val new_frames = frame::konts
  val baddr = allocBind(v, new_time)
  val new_env = c_env + (v -> baddr)
  val argvs = atomicEval(ae, env, bstore)
  val new_store = bstore.update(baddr, argvs)
  val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
  NDState(body, new_env, new_store, new_frames, new_time, new_ndk)
\end{lstlisting}

By atomically evaluating @f@, we obtain a set of closures.
The transition is deterministic to the first element of the closure set.
So we just prepare new environment, new store, and new frame list only for the
first closure in that set.
A new non-deterministic continuation @new_ndk@ is also constructed,
which contains the rest of closures, the values of argument @ae@, the store,
the time and the new frame list.
We use the store before updating because for different closures they may form
different binding addresses @baddr@.
We also use the new frames, because all the closures at this fork point share
the same stack and return point.

In the second case shown in below we will see how this @NDCont@ controls
the non-determinism.

\begin{lstlisting}
case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
  konts match {
    case Nil => ndk match {
      case NDCont(Nil, _, _, _, _)::ndk => 
        NDState(ae, env, bstore, konts, time, ndk) /* transfer to the next fork point */
      case NDCont(cls, argvs, bstore, time, frames)::ndk => 
        val Clos(Lam(v, body), c_env) = cls.head
        val baddr = allocBind(v, time)
        val new_env = c_env + (v -> baddr)
        val new_store = bstore.update(baddr, argvs)
        val new_ndk = NDCont(cls.tail, argvs, bstore, tile, frames)::ndk
        /* resume the fork point with the next closure */
        NDState(body, new_env, new_store, frames, time, new_ndk) 
    }
    case Frame(x, e, f_env)::konts =>
      val baddr = allocBind(x, newTime)
      val new_env = f_env + (x -> baddr)
      val new_store = bstore.update(baddr, aeval(ae, env, bstore))
      NDState(e, new_env, new_store, konts, newTime, ndk) /* normal return */
  }
\end{lstlisting}

If the normal continuation @konts@ is an empty list, then it means we have reached
the halt of computation; otherwise, we should return the values of @ae@ to its
caller which is contained in the top frame of stack.

But, since we have added non-determinism into state, an empty list of frames means
we have the halt of \textit{one computation path}, and we should determine the next state
by @NDCont@.
So we have a pattern matching on @ndk@: 1) If the closure set is an empty set,
then we had tried all the closures of this fork point and should move the the
next fork point. 2) Otherwise, we pop a closure from the set, then build new
environment and new store for the body expression of the closure;
we use the same frame list and time that are copied from the fork point rather
than current one.
The non-determinism continuation @new_ndk@ is also updated by removal of that closure.

\begin{lstlisting}
def drive(nds: NDState, seen: Set[State]): Set[State] = {
  nds match {
    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
    case nds =>
      val s = nds.toState
      if (seen.contains(s)) drive(step(nds), seen)
      else drive(step(nds), seen + s)
  }
}
\end{lstlisting}

The function @drive@ is also changed: there is no work list anymore, because all the
potentially unexplored states are embedded in to the continuation for
non-determinism.
The termination of the analysis is when we reach a @NDState@ that the expression
is atomic, and both the normal continuation and non-determinism continuation are
empty list. This corresponds to the case that @todo@ list is empty.

Now, we obtain a linearized abstract abstract machine. If we image that the classical
AAM explores a graph of reachable states, then the linearized AAM flatten this
graph.

@@

Fusing is a tramsformation that combines function @step@ and @drive@ to one.
The fused function @drive_step@ basically formed by merging function
@step@ into function @drive@.
@drive_step@ takes a @NDState@ and a set of explored states as argument,
and returns a set of reachable states once terminates.

\begin{lstlisting}
def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
  nds match {
    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
    case nds =>
      val newTime = tick(nds)
      val new_ndstate = nds match {
        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) =>
          ......
        case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) =>
          ......
      }
      val s = nds.toState
      if (seen.contains(s)) drive_step(new_ndstate, seen)
      else drive_step(new_ndstate, seen + s)
  }
}
\end{lstlisting}

Now we have a single function to perform both abstract evaluation and intermediate
states collecting. Given @inject(e)@ as the inital state and empty set as
initial reached set, we can easily define the entrance function @analyze@.

\begin{lstlisting}
def analyze(e: Expr): Set[State] = drive_step(inject(e), Set())
\end{lstlisting}

@@

In last section, the fused AAM combines all the functions into one and
looks like a ``big-step'' interpreter, although it still has the machine
state representation inside. In the disentangling transformation, we

\begin{lstlisting}
\end{lstlisting}

@@

Defunctionalization and refunctionalization build connections between abstract
machines and interpreters.

For deterministic languages, a reduction semantics is a structural operational
semantics in continuation style, where the reduction context is a
defunctionalized continuation.
Unfortunately, the abstract abstract machines is non-deterministic, where
the non-determinism is handled by a work list.
We can turn this explicit non-determinism into the reduction context of
AAM.

@@

\todo{PDCFA, AAC}
